---
output: 
  github_document:
    self_contained: no
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
    echo = TRUE,
    fig.align="center",
    fig.width = 6,
    fig.height = 4)
```

<!-- README.md is generated from README.Rmd. Please edit that file -->

# hyd1d

[![pipeline status](https://git.aqualogy.de/arnd/hyd1d/badges/master/pipeline.svg)](https://git.aqualogy.de/arnd/hyd1d/commits/master)

The R package **hyd1d** is designed to compute 1-dimensional water level 
information along German federal waterways Elbe and Rhine.

## Installation

**hyd1d** is not currently available from CRAN, but you can install the 
development version from BfG's gitbucket server with:

```{r install_git, eval = FALSE}
install.packages("devtools")
library(devtools)
devtools::install_git("git://apps.bafg.de/gitbucket/webera/hyd1d.git")
```

## Usage

The package **hyd1d** is build around the S4-class `WaterLevelDataFrame`. To 
compute and visualize 1-dimensional water level information an object of class 
`WaterLevelDataFrame` has to be initialized. Various functions included in 
**hyd1d** use these objects and compute water levels stored in the column `w`.

```{r library, error = FALSE, message = FALSE, warning = FALSE, include = FALSE}
# standard library path for the installed local packages
R_version <- paste(sep = ".", R.Version()$major, R.Version()$minor)
lib <- paste0("~/R/", R_version, "/")
library(hyd1d, lib.loc = lib)

# set english locale to produce english plot labels
Sys.setlocale(category = "LC_MESSAGES", locale = "en_US")
```

```{r usage, eval = TRUE, error = FALSE, warning = FALSE, message = FALSE}
# load the package
library(hyd1d)

# initialize a WaterLevelDataFrame
wldf <- WaterLevelDataFrame(river   = "Elbe",
                            time    = as.POSIXct("2016-12-21"),
                            station = seq(257, 262, 0.1))

# compute a water level
wldf <- waterLevel(wldf, TRUE)

# and plot it
plotShiny(wldf, TRUE, TRUE, TRUE)
```
