---
title: "hyd1d"
author: "Arnd Weber"
date: "`r Sys.Date()`"
output:
  html_vignette:
    toc: yes
    toc_depth: 3
bibliography: ../inst/REFERENCES.bib
csl: apa_modified.csl
vignette: |
  %\VignetteIndexEntry{hyd1d} 
  %\VignetteEngine{knitr::rmarkdown} 
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    fig.align="center",
    fig.width = 7,
    fig.height = 4, 
    root.dir = "vignettes"
)
```

<!-- Indent in html version TOC
     obtained from: 
     https://stackoverflow.com/questions/46201753/rmarkdown-indentation-of-toc-items-in-html-output
     -->

<script>
$(document).ready(function() {
  $items = $('div#TOC li');
  $items.each(function(idx) {
    num_ul = $(this).parentsUntil('#TOC').length;
    $(this).css({'padding-left': num_ul * 5px, 'box-shadow': none});
  });

});
</script>

<!-- numbering for figures and tables-->

```{r captions, echo = FALSE, error = FALSE, warning = FALSE, message = FALSE, include = FALSE}
require(hyd1d)
require(plotrix)

# set english locale to produce english plot labels
Sys.setlocale(category = "LC_MESSAGES", locale = "en_US.utf8")

# Determine the output format of the document
outputFormat <- knitr::opts_knit$get("rmarkdown.pandoc.to")

# Figure and Table Caption Numbering, for HTML do it manually
capTabNo <- 1
capFigNo <- 1

# Function to add the Table Number
capTab <- function(x){
    if(outputFormat == 'html'){
        x <- paste0("**Tab. ", capTabNo, "**: ", x)
        capTabNo <<- capTabNo + 1
    }
    return(x)
}

# Function to add the Figure Number
capFig <- function(x){
    if(outputFormat == 'html'){
        x <- paste0("**Fig. ", capFigNo, "**: ", x)
        capFigNo <<- capFigNo + 1
    }
    return(x)
}
```

<BR>

# Purpose

**hyd1d** is an R package that provides a S4-class and several functions to 
compute 1-dimensional water levels along the German federal waterways Elbe and 
Rhine.

<BR>

# Use

## Installation

The package **hyd1d** is not available from CRAN at the moment. To install the 
recent developmental version from the GitBucket server of the Federal Institute 
of Hydrology (BfG) execute the following commands: 

```{r install_git, eval = FALSE}
install.packages("devtools")
library(devtools)
devtools::install_git("git://apps.bafg.de/gitbucket/webera/hyd1d.git")
```

Afterwards **hyd1d** can be loaded like every other R package with the following 
command: 

```{r library, eval = TRUE, echo = TRUE, error = FALSE, warning = FALSE, message = FALSE}
library(hyd1d)
```

## S4-class WaterLevelDataFrame

Basis of all water level computations with **hyd1d** is the S4-class 
[`WaterLevelDataFrame`](reference/WaterLevelDataFrame-class.html). To compute 
water levels with one of the `waterLevel...`-functions, a `WaterLevelDataFrame`
has to be initialized with the homonymous 
[`WaterLevelDataFrame()`](reference/WaterLevelDataFrame.html)-function: 

```{r wldf, eval = TRUE}
wldf <- WaterLevelDataFrame(river   = "Elbe",
                            time    = as.POSIXct("2016-12-21"),
                            station = seq(257, 262, 0.1))
```

The minimal required information to initialize a `WaterLevelDataFrame` are the 
the function arguments `river`, `time` and at least one of the two possible 
station arguments (`station` or `station_int`). With this information an object 
of class `WaterLevelDataFrame` can be created, which has the following 
structure:

```{r structure, eval = TRUE}
str(wldf)

summary(wldf)

```

The actual water level information is stored in the S4-slot `.Data`, which is in 
fact a `data.frame`with the columns `station`, `station_int` and `w`. The 
columns `station` and `station_int` contain a stationing information, which 
corresponds to the official stationing of the waterway and navigation agency 
(Wasserstraßen- und Schifffahrtsverwaltung). The stationing information is 
duplicated to enable database joins with GIS data through the `integer`-type 
column `station_int`. The column `w` contains the actual 
water level in the height reference system DHHN92 (German height reference 
system 1992), but is usually empty after initialization and gets filled 
throught the application of one of the `waterLevel...`-functions.

For the application of `waterLevel...`-functions information stored in the 
S4-slots `river` and where appropriate `time` is essential. They enable a 
distinct localization of the stationing along the rivers Elbe and Rhine and a 
determination of the time of the water level computation. The other slots of an 
object of class `WaterLevelDataFrame` are filled during the water level 
computation and contain partial results needed to visualize the results 
(`gauging_station`) or have an informative character (
`gauging_stations_missing`, `comment`).

## Computation of water levels

### waterLevel

The most advanced function to interpolate FLYS3 water levels with local gauging 
data is implemented in the [`waterLevel`](reference/waterLevel.html)-function. 
This function uses package-internal gauging data from the dataset 
[`df.gauging_data`](reference/df.gauging_data.html), which contains daily-
averaged gauging data since 1990-01-01. Therefore 
[`waterLevel`](reference/waterLevel.html) can be applied for the time period 
between 1990-01-01 and yesterday.

After the [initialization](articles/hyd1d.html#s4-klasse-waterleveldataframe) of 
a `WaterLevelDataFrame` the application is very simple:

```{r waterlevel, eval = TRUE}
wldf <- waterLevel(wldf)
summary(wldf)

```

And if you want to visualize the results using [`plotShiny`](reference/plotShiny.html)
the additional argument `shiny = TRUE` has to be used. Thereby the columns 
`section`, `weight_x` and `weight_y` get created in the `.Data`-Slot, which are 
required for the visualization.

```{r figure1, fig.show = 'asis', fig.cap = capFig("Interpolated water level, computation-relevant stationary FLYS3 water levels (**0.5MQ**, **a** and **0.75MQ**) and gauge heigth of 2016-12-21 at River Elbe between Rosslau and Dessau."), eval = TRUE}
wldf <- waterLevel(wldf, shiny = TRUE)
summary(wldf)

xlim_min <- 257
xlim_max <- 263
{
    plotShiny(wldf, TRUE, TRUE, TRUE, xlim = c(xlim_min, xlim_max),
              xlab = "river station (km)", ylab = "elevation (m a.s.l. (DHHN92))")
    legend("topright", 
           col = c("darkblue", "darkblue", "darkblue", "red", "black"), 
           pch = c(21, 0, NA, NA, NA), 
           pt.bg = c("darkblue", "darkblue", NA, NA, NA), 
           pt.cex = c(1, 1, NA, NA, NA), 
           lty = c(0, 0, 1, 1, 1), 
           lwd = c(0, 0, 1, 0.6, 0.6), 
           legend = c("gauge height", "gauge weight", "waterLevel", 
                      "upper FLYS w.l.", "lower FLYS w.l."), 
           text.col = c(1, "darkblue", 1, 1, 1), 
           cex = 0.7, bty = "n")
}
```

### waterLevelPegelonline

The way how the [`waterLevelPegelonline`](reference/waterLevel.html)-function
computes a water level is equivalent to the [`waterLevel`](reference/waterLevel.html)-
function. Just the data source for the gauging data is different, since it does 
not use package-internal data, but online data provided by <https://pegelonline.wsv.de>.
Because data provided by PEGELONLINE are available for only 30 days backwards, 
the application of this function is limited to recent time periods, but with a 
high temporal resolution.

```{r figure2, fig.show = 'asis', fig.cap = capFig(paste0("Interpolated water level, computation-relevant stationary FLYS3 water levels (**0.5MQ**, **a** and **0.75MQ**) and gauge heigth of ", strftime(Sys.time() - 3600, format = "%Y-%m-%d %H:%M")," at River Elbe between Rosslau and Dessau.")), eval = TRUE}
# one hour ago
time <- as.POSIXct(Sys.time() - 3600)

# initialize a WaterLevelDataFrame
wldf <- WaterLevelDataFrame(river   = "Elbe",
                            time    = time,
                            station = seq(257, 262, 0.1))

# compute w
wldf <- waterLevelPegelonline(wldf, shiny = TRUE)
summary(wldf)

# and plot the results
{
    plotShiny(wldf, TRUE, TRUE, TRUE, xlim = c(xlim_min, xlim_max),
              xlab = "river station (km)", ylab = "elevation (m a.s.l. (DHHN92))")
    legend("topright", 
           col = c("darkblue", "darkblue", "darkblue", "red", "black"), 
           pch = c(21, 0, NA, NA, NA), 
           pt.bg = c("darkblue", "darkblue", NA, NA, NA), 
           pt.cex = c(1, 1, NA, NA, NA), 
           lty = c(0, 0, 1, 1, 1), 
           lwd = c(0, 0, 1, 0.6, 0.6), 
           legend = c("gauge height", "gauge weight", "waterLevel", 
                      "upper FLYS w.l.", "lower FLYS w.l."), 
           text.col = c(1, "darkblue", 1, 1, 1), 
           cex = 0.7, bty = "n")
}

```

### waterLevelFlood1 & waterLevelFlood2

To compare the newly developed functions [`waterLevel`](reference/waterLevel.html) 
and [`waterLevelPegelonline`](reference/waterLevel.html) to existing computation
methods, the functions [`waterLevelFlood1`](reference/waterLevelFlood1.html) and
[`waterLevelFlood2`](reference/waterLevelFlood2.html) have been implemented. 
These functions compute water levels according to the Flood1- and Flood2-methods
of the modelling environment [INFORM](http://www.bafg.de/DE/08_Ref/U2/02_analyse/01_INFORM/inform.html) 
[@rosenzweig_inform_2011]. They either shift the reference water level **MQ**
vertically, so that it intersects with the gauge height at a selected reference
gauging station, or linearily interpolate water levels with neighboring gauging
stations.

```{r figure3, fig.show = 'asis', fig.cap = capFig("Water levels computed after the Flood1-method with the reference gauges Rosslau (wldf1) and Dessau (wldf2) and the Flood2-method at the 2016-12-21 at River Elbe between Rosslau and Dessau."), eval = TRUE}
wldf <- WaterLevelDataFrame(river   = "Elbe",
                            time    = as.POSIXct("2016-12-21"),
                            station = seq(257, 262, 0.1))

wldf1 <- waterLevelFlood1(wldf, "ROSSLAU", shiny = TRUE)
summary(wldf1)

wldf2 <- waterLevelFlood1(wldf, "DESSAU", shiny = TRUE)
summary(wldf2)

wldf3 <- waterLevelFlood2(wldf)
summary(wldf3)

df.gs2 <- getGaugingStations(wldf2)

{
    plotShiny(wldf1, FALSE, FALSE, FALSE, xlim = c(xlim_min, xlim_max),
              xlab = "river station (km)", ylab = "elevation (m a.s.l. (DHHN92))")
    lines(wldf2$station, wldf2$w, col = "darkblue", lty = 2)
    lines(wldf3$station, wldf3$w, col = "red", lty = 2)
    abline(v = df.gs2$km_qps, lty = 3, lwd = 0.5)
    points(df.gs2$km_qps, df.gs2$wl, pch=21, col="darkblue", bg="darkblue")
    boxed.labels(df.gs2$km_qps, 55.4, df.gs2$gauging_station, bg="white", 
                 srt = 90, border = FALSE, xpad = 4, ypad = 0.7, cex = 0.7)
    legend("topright", 
           col = c("darkblue", "darkblue", "darkblue", "red"), 
           pch = c(21, NA, NA, NA), 
           pt.bg = c("darkblue", NA, NA, NA), 
           pt.cex = c(1, NA, NA, NA), 
           lty = c(0, 1, 2, 2), 
           lwd = c(0, 1, 1, 1), 
           legend = c("gauge height", "wldf1", "wldf2", "wldf3"), 
           cex = 0.7, bty = "n")
}
```

### waterLevelFlys3InterpolateY

To compare the newly developed functions [`waterLevel`](reference/waterLevel.html) 
and [`waterLevelPegelonline`](reference/waterLevel.html) to existing computation
methods of [FLYS3](http://www.bafg.de/DE/08_Ref/M2/03_Fliessgewmod/01_FLYS/flys_node.html),
the function [`waterLevelFlys3InterpolateY`](reference/waterLevelFlys3InterpolateY.html) 
has been implemented. This function computes a water level according to the 
method implemented in the W-INFO-module of FLYS3. This function determines the 
relative position of the gauge height at a reference gauge to the two 
surrounding FLYS3 water levels and uses this weight for a longitudinal 
interpolation between both water levels.

```{r figure4, fig.show = 'asis', fig.cap = capFig("Water levels after FLYS3 with the reference gauge Rosslau at the 2016-12-21 at River Elbe between Rosslau and Dessau."), eval = TRUE}
wldf <- waterLevelFlys3InterpolateY(wldf, "ROSSLAU", shiny = TRUE)
summary(wldf)

{
    plotShiny(wldf, TRUE, TRUE, TRUE, xlim = c(xlim_min, xlim_max),
              xlab = "river station (km)", ylab = "elevation (m a.s.l. (DHHN92))")
    abline(v = df.gs2$km_qps, lty = 3, lwd = 0.5)
    points(df.gs2$km_qps, df.gs2$wl, pch=21, col="darkblue", bg="darkblue")
    boxed.labels(df.gs2$km_qps, 55.4, df.gs2$gauging_station, bg="white", 
                 srt = 90, border = FALSE, xpad = 4, ypad = 0.7, cex = 0.7)
    legend("topright", 
           col = c("darkblue", "darkblue", "darkblue", "red", "black"), 
           pch = c(21, 0, NA, NA, NA), 
           pt.bg = c("darkblue", "darkblue", NA, NA, NA), 
           pt.cex = c(1, 1, NA, NA, NA), 
           lty = c(0, 0, 1, 1, 1), 
           lwd = c(0, 0, 1, 0.6, 0.6), 
           legend = c("gauge height", "gauge weight", "waterLevel", 
                      "upper FLYS w.l.", "lower FLYS w.l."), 
           text.col = c(1, "darkblue", 1, 1, 1), 
           cex = 0.7, bty = "n")
}
```

### waterLevelFlys3...

All other `waterLevelFlys3...`-functions (
[`waterLevelFlys3`](reference/waterLevelFlys3.html), 
[`waterLevelFlys3Seq`](reference/waterLevelFlys3.html) and [`waterLevelFlys3InterpolateX`](reference/waterLevelFlys3InterpolateX.html)) 
serve exclusively for the preparation and querying of stationary FLYS3 water 
levels. They can be used to extract water levels from the dataset 
[`df.flys_data`](reference/df.flys_data.html) and interpolate the water levels 
along the x-axis linearily, but without modifying the datasets contents. These
functionalities are needed for **all**  `waterLevel...`-functions described in 
this vignette and are mentioned here for completeness.

<BR>

# References

---
nocite: | 
  @wsv_pegelonline_2018
  @bundesanstalt_fur_gewasserkunde_flys_2016
...

